\emph{icepack}: a novel glacier flow modeling package

### 2019-03-26 icepack.tex
Started writing about inverse problems

>>>>

Icepack includes a set of routines for estimating the basal friction or rheology coefficients from observational data.
The class \texttt{InverseProblem} represents the specification of the inverse problem.
To specify an inverse problem, the user must provide:
\begin{itemize}
    \item the model object and the method that solves the diagnostic equation,
    \item the objective and regularization functionals,
    \item the observed field and the name of the argument to the diagnostic solver,
    \item an initial guess for the field to be estimated and the name of the argument to the diagnostic solver,
    \item extra data passed to the diagnostic solver such as boundary conditions.
\end{itemize}
At present, this class assumes that the observed state is always the ice velocity, but in principle the same design would suffice for more complicated inverse problems.
More importantly, the inverse problem is flexible enough to account for the fact that we might have substituted in our own parameterization for the rheology or friction coefficient.
For example, nearly all studies in the literature use an alternate parameterization of these coefficients in terms of some auxiliary field in order to guarantee positivity \textcolor{red}{citations}.
This information is passed to the inverse problem by specifying the names of the keyword arguments for the observed state and the field we wish to observe.

The class \texttt{InverseSolver} is an abstract interface for classes that actually solve the inverse problem.
There are three implementations of this interface in icepack: \texttt{GradientDescentSolver}, \texttt{BFGSSolver}, and \texttt{GaussNewtonSolver}.
All of these classes are based around the general idea of first computing a search direction and then performing a line search in every iteration.
They differ in how the search direction is computed.
The gradient descent solver uses the search direction
\begin{equation}
    \phi_k = -M^{-1}\ud J(\theta_k)
\end{equation}
where $M$ is the finite element mass matrix.
Gradient descent is a popular choice because the objective functional is always decreasing along this search direction.
However, the search direction can be poorly scaled to the physical dimensions of the problem at hand.
This method can be very expensive and brittle in the initial iterations and often takes many steps to converge.
The BFGS method uses the past $m$ iterations of the algorithm to compute a low-rank approximation to the inverse of the objective functional's Hessian matrix; see \citet{nocedal2006numerical} for a more in-depth discussion.
The BFGS method converges faster than gradient descent but suffers from many of the same brittleness issues in the initial iterations before it has accumulated enough history.
Finally, the Gauss-Newton solver defines an approximation to the ``first-order'' part of the objective functional Hessian.
Each iteration of Gauss-Newton is more expensive than that of BFGS or gradient descent because it requires the solution of a more complex linear system than just the mass matrix.
The Gauss-Newton method converges fastest by far in virtually every test case we have found, in some instances by up to factor of 50.

The derivative of the objective functional with respect to the unknown parameter is calculated using the symbolic differentiation features of Firedrake.
The user does not need to provide any routines for the derivatives, only the symbolic form of the error metric and the regularization functional.
The model object is responsible for providing the symbolic form of the action functional.

The bootstrapping approach used to calculate higher-order solutions to the hybrid model from lower-order solutions rather than from a cold initial guess can also be used for inverse problems.
First, we estimate the unknown parameter using only degree-1 vertical basis functions for the velocity field, which captures most of the spatial variability of the unknown parameter.
The resulting coefficient is then used as the initial guess with higher degree vertical basis functions for the velocity \textcolor{red}{benchmark}.
